Performance for the masses: Experiments with A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Last changes

So, the rest of the paper is organized as follows: next we present the
state of the art in web-based distributed
computational systems along with attempts to predict and model their
behavior in Section \ref{sec:soa}. A brief presentation of the
methodology used and the results of the experiments are presented in
%Pablo: mention the problems used in the first experiments
Section \ref{sec:exp1}, including new results using web workers on the
same problem and on a new problem, a modified Rastrigin (F15) function. 
%Pablo: Rastrigin is not the owner of the function
%Mario: In many places is called  Rastrigin's function because it was proposed by him.
Finally, conclusions and future lines of work are presented in Section
\ref{sec:conclusion}.

@@

\subsection{A hard optimization problem: Rastrigin's function}
\label{sec:rastrigin}

The fact that the solution of this particular problem was found in a
few seconds and restarts were necessary yielded interesting results,
but we needed to work on a harder problem, that is why created a new
client version that optimized Rastrigin function, a multimodal
floating-point problem used as a benchmark in evolutionary
algorithms. The set up for this function, and a comparison of the
baseline speed for JavaScript in an standalone experiment and other
languages is shown in \cite{2016arXiv160101607Manom}, where the details of
the implementation have also been extensively described.

@@

The results for this new experiment have been extremely interesting
and quite different from the previous one. The most important
characteristic that set this experiment apart from the previous one is
the fact that it did not find the solution for the time it was
running, instead lowering, little by little, the fitness until it was
disconnected. The fitness of chromosomes sent by clients managed to
surpass the 9000 value by the end of the; and although initially there
was a high number of new nodes, this number decreases with time.

This is in part due to the fact that the presence of nodes is not
constant, as shown in Figure \ref{fig:ips:rastrigin}, which plots the
number of unique IPs participating in the experiment per minute (left)
and hour (right). The two graphs have the same shape: At the beginning there are very few or only
one, but later one there are peaks of almost 30 unique machine a
minute; these peaks appear again when new announcements are made in
Twitter or any other social networks, to go down to a few remaining
nodes later on. It is interesting to note that, at the most, more than
100 unique nodes were participating in the experiment during one hour;
also that there are sustained peaks during a day of 6-7 computers
every minute. This is due to the fact that the computation is
asynchronous, but also to the fact that different computers stay in
the simulation for a different period of time. This period has been
plotted in Figure \ref{fig:fit:rastrigin}, that plots the number of
PUTs per node, ranked by number, along with a fit to the two
distributions we have used before, GEV and Weibull; the parameters of
this fit are shown in Table \ref{tab:puts:ww:f15}. Unlike the previous
experiments, in which GEV fit was relatively good, in this case it is
quite obviously not. However, the fit to the Weibull distribution,
shown at the left of the image, is relatively good. As should be
expected, the shape and scale are different to the one in the previous
experiments with the Trap function, however, the shape is quite
similar (0.40 vs 0.69), being in both cases less than one, which
indicates a decreasing distribution of the same kind. In fact, the
difference in the scale parameter indicates that the difference
between the time devoted by a volunteer and the next is bigger in this
case. This is probably due to the fact that the duration of the
experiment is bigger, allowing for bigger differences in
behavior. While the first in the Trap function might have devoted a
few minutes and the second a few seconds less, the difference between
the first and second in this experiment might be in the scale of
hours.

In general, this experiment confirms what was found in the others in
several aspects: first, there is a regularity in the distribution of the
time assigned by volunteers to it, second, there is a good and certain
amount of volunteers that will become part of the distributed
computer, in some cases up to one hundred; also, in some cases there
will be users that will spend several hours, two orders of magnitude
more generations in this case than in the previous one. There is a
clear influence of the time to solution: a short time means that every
experiment will only gather a few volunteers, but also that they will
stay only for a short time. A long time will make users stay for
longer, but if it takes too long, they will eventually lose interest.

>>>>

So, the rest of the paper is organized as follows: next we present the
state of the art in web-based distributed
computational systems along with attempts to predict and model their
behavior in Section \ref{sec:soa}. A brief presentation of the
methodology used and the results of the experiments are presented in
%Pablo: mention the problems used in the first experiments
Section \ref{sec:exp1}, including new results using web workers on the
same problem and on a new problem, a modified Rastrigin's function. 
%Pablo: Rastrigin is not the owner of the function
%Mario: In many places is called  Rastrigin's function because it was proposed by him.
Finally, conclusions and future lines of work are presented in Section
\ref{sec:conclusion}.

@@

\subsection{A hard optimization problem: A Shifted Rotated Rastrigin's function}
\label{sec:rastrigin}

The fact that the solution of this particular problem was found in a
few seconds and restarts were necessary yielded interesting results,
but we needed to work on a harder problem, that is why created a new
client version that optimized a harder variation of the Rastrigin's function, 
used as a benchmark in evolutionary
algorithms. The set up for this function, and a comparison of the
baseline speed for JavaScript in an standalone experiment and other
languages is shown in \cite{2016arXiv160101607Manom}, where the details of
the implementation have also been extensively described.

@@

The results for this new experiment have been extremely interesting
and quite different from the others. The most important
characteristic that set this experiment apart is
the fact that it did not find the solution for the time it was
running, instead lowering, little by little, the fitness until it was
disconnected. The fitness of chromosomes sent by clients managed to
surpass the 9000 value by the end of the execution; and although initially there
was a high number of new nodes, this number decreased with time.

This is in part due to the fact that the presence of nodes is not
constant, as shown in Figure \ref{fig:ips:rastrigin}, which plots the
number of unique IPs participating in the experiment per minute (left)
and hour (right). The two graphs have the same shape: At the beginning there are very few or only
one, but later one there are peaks of almost 30 unique machines a
minute; these peaks appear again when new announcements are made in
Twitter or any other social networks, to go down to a few remaining
nodes later on. It is interesting to note that, at the most, more than
100 unique nodes were participating in the experiment during one hour;
also that there are sustained peaks during a day of 6-7 computers
every minute. This is due to the fact that the computation is
asynchronous, but also to the fact that different computers stay in
the simulation for a different period of time. This period has been
plotted in Figure \ref{fig:fit:rastrigin}, that plots the number of
PUTs per node, ranked by number, along with a fit to the two
distributions we have used before, GEV and Weibull; the parameters of
this fit are shown in Table \ref{tab:puts:ww:f15}. Unlike the previous
experiments, in which GEV fit was relatively good, in this case it is
quite obviously not. However, the fit to the Weibull distribution,
shown at the left of the image, is relatively good. As should be
expected, the shape and scale are different to the one in the previous
experiments with the Trap function, however, the shape is quite
similar (0.40 vs 0.69), being in both cases less than one, which
indicates a decreasing distribution of the same kind. In fact, the
difference in the scale parameter indicates that the difference
between the time devoted by a volunteer and the next is bigger in this
case. This is probably due to the fact that the duration of the
experiment is bigger, allowing for bigger differences in
behavior. While the first in the Trap function might have devoted a
few minutes and the second a few seconds less, the difference between
the first and second in this experiment might be in the scale of
hours.

In general, this experiment confirms what was found in the others in
several aspects: first, there is a regularity in the distribution of the
time assigned by volunteers to it, second, there is a good and certain
amount of volunteers that will become part of the distributed
computer, in some cases up to one hundred; also, in some cases there
will be users that will spend several hours, two orders of magnitude
more generations in this case than in the previous one.
%Mario: than in the previous one. Are we refereing to the Wˆ2 version?
There is a clear influence of the time to solution: a short time means that every
experiment will only gather a few volunteers, but also that they will
stay only for a short time. A long time will make users stay for
longer, but if it takes too long, they will eventually lose interest.
